---
title: "LAPD data exploration"
author: "Lael Hebert"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r}
library(dplyr)
library(readr)
library(lubridate)
library(stringr)
library(tidyr)
```

## R Load data

Start by loading all the data into R with good column types. Fix column names.

```{r}
census_data <- read_csv("Census_Data_by_Neighborhood_Council.csv", 
    col_types = cols(Hispanic_pop = col_double(), 
        Owner_occ = col_skip(), Renter_occ = col_skip()))
census_data <- rename(census_data, neighborhood = NC_Name, Total_pop = "Total Population")

reporting_districts <- read_csv("LAPD_Reporting_Districts.csv", 
    col_types = cols(AREA = col_skip(), OBJECTID = col_skip(), 
        PERIMETER = col_skip(), PREC = col_character(), 
        REPDIST = col_character(), Shape__Area = col_skip(), 
        Shape__Length = col_skip()))
reporting_districts <- rename(reporting_districts, reporting_district = REPDIST, division_number = PREC, division = APREC, bureau = BUREAU, basic_car = BASICCAR)

stops_data <- read_csv("Vehicle_and_Pedestrian_Stop_Data_2010_to_Present.csv", 
    col_types = cols(`Form Reference Number` = col_character(), 
        `Officer 1 Serial Number` = col_character(), 
        `Reporting District` = col_character(), 
        `Stop Date` = col_date(format = "%m/%d/%Y"), 
        `Stop Number` = col_character(), 
        `Stop Time` = col_time(format = "%H:%M")))
names(stops_data) = c("stop_number", "form_reference_number", "gender_code", "ethnicity_code", "ethnicity", "date", "time", "officer_1_serial", "officer_1_division_num", "division_1", "officer_2_serial", "Officer_2_division_num", "division_2", "reporting_district", "stop_type", "post_stop_activity")

division_populations_all <- read_csv("Division_populations.csv") 

division_pop <- read_csv("Division_populations.csv", 
    col_types = cols(Asian_pop = col_skip(), 
        Black_pop = col_skip(), Hispanic_pop = col_skip(), 
        Other_pop = col_skip(), White_pop = col_skip(), 
        total_pop = col_skip()))
division_pop <- rename(division_pop, WHITE = white_percent, BLACK = black_percent, ASIAN = asian_percent, HISPANIC = hispanic_percent, OTHER = other_percent)


```
Clean up the stops data: 
Add a year column. 
If the reporting district starts with 0, remove the 0.
Join the stops data with the reporting districts data to get the division and bureau where the stop took place. 
Rename the new columns in the stops data.

```{r}
stops_data <- stops_data %>%
  mutate(year = year(date)) %>%
  mutate(reporting_district = ifelse(grepl(pattern = "^0", reporting_district), 
                                 str_sub(reporting_district, start = 2), reporting_district)) %>%
  left_join(reporting_districts, "reporting_district") %>%
  rename(stop_division_number = division_number, stop_division = division, stop_bureau = bureau, stop_basic_car = basic_car)


```

The population data is missing from the hispanic column in the census data. Fill it in by subtracting the popultions of the other ethnicities from the total population. Calculate the total population of the city. Then created a vector of the percentages of each ethnicity city-wide. 


```{r}
census_data$Hispanic_pop <- census_data$Total_pop - 
  census_data$White_pop -
  census_data$Black_pop -
  census_data$Ameri_es_pop -
  census_data$Asian_pop -
  census_data$Hawn_pi_pop -
  census_data$Other_pop -
  census_data$Multi_pop

city_pop_total <- sum(census_data$Total_pop)
city_wide_percentages <- census_data[,3:10] %>%
  summarise_each(funs(sum))/city_pop_total

city_pop_total
city_wide_percentages



```

According to the [Census.gov website](https://www.census.gov/quickfacts/fact/table/losangelescitycalifornia,losangelescountycalifornia/POP010210), Los Angelos had a population of 3.79 million people in 2010. This is very close to the 3.73 million reported in our census data set from the neighborhood council. The percentages by each ethnicity also match well. It looks like our calculation for the hispanic population will work.

Now, let's combine all ethnicities that represent less than 2% of the population as "other." We'll change the ethnicity column in stops_data but leave the ethnicity_code column alone in case we need the original values later. 
```{r}
unique(stops_data$ethnicity)
stops_data <- stops_data %>% 
  mutate(ethnicity = ifelse(ethnicity %in% c("OTHER", "AMERICAN INDIAN", "MULTI-DESCENTS"), "OTHER", ethnicity))
unique(stops_data$ethnicity)


```
Add a column to the census_data called All_other_pop with the sum of the populations of ethnic groups with less than 2% of the population.  
```{r}
census_data <- census_data %>% mutate(
  All_other_pop = 
    Ameri_es_pop +
    Hawn_pi_pop +
    Other_pop + 
    Multi_pop
)
```
Create some data sets: one with stop percentage by ethnicity for each division for all years. The second will have the same data broken out by year.  For each data set, add data with the ethnicity data by division populations for the neighborhood.  
```{r}
stops_summary_all_years <- stops_data %>% 
  filter(!is.na(stop_division)) %>%
  group_by(stop_division, ethnicity) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(location = "stops") %>%
  rename(division = stop_division)

stops_summary_ethnicity_totals_all_years <- stops_data %>% 
  filter(!is.na(stop_division)) %>%
  group_by(ethnicity) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(location = "stops") %>% 
  mutate(division = "CITY TOTAL")

stops_summary_ethnicity_totals_all_years <- stops_summary_ethnicity_totals_all_years[c(5,1,2,3,4)]
stops_summary_all_years <- bind_rows(stops_summary_all_years, stops_summary_ethnicity_totals_all_years)

stops_summary_by_year <- stops_data %>% 
  filter(!is.na(stop_division)) %>%
  group_by(stop_division, year, ethnicity) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(location = "stops") %>%
  rename(division = stop_division)

stops_summary_ethnicity_totals_by_year <- stops_data %>% 
  filter(!is.na(stop_division), year != 1900) %>%
  group_by(year, ethnicity) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(location = "stops") %>%
  mutate(division = "CITY TOTAL")

stops_summary_ethnicity_totals_by_year <- stops_summary_ethnicity_totals_by_year[c(6,1,2,3,4,5)]
stops_summary_by_year <- bind_rows(stops_summary_by_year, stops_summary_ethnicity_totals_by_year)

#division_populations_all so that it is in the same form as stops_summary. Call it division_pop_gathered
division_counts <- division_populations_all[1:7] %>% 
  rename(WHITE = White_pop, BLACK = Black_pop, ASIAN = Asian_pop, HISPANIC = Hispanic_pop, OTHER = Other_pop) %>%
  gather("ethnicity", "ethnicity_count", 2:7) 
division_percents <- division_populations_all %>%
  select(1,8:12) %>%
  rename(WHITE = white_percent, BLACK = black_percent, ASIAN = asian_percent, HISPANIC = hispanic_percent, OTHER = other_percent)  %>%
  gather("ethnicity", "ethnicity_percent", 2:6)

division_pop_gathered <- left_join(division_counts,division_percents) %>%
  mutate(location = "population")

stops_summary_all_years_with_pop <- bind_rows(stops_summary_all_years, division_pop_gathered)
stops_summary_by_year_with_pop <- bind_rows(stops_summary_by_year, mutate(division_pop_gathered, year = 2010))

write_csv(stops_summary_all_years_with_pop, path = "C:/Users/lahebert.REDMOND/Documents/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/stops_summary_all_years_with_pop.csv")
write_csv(stops_summary_by_year_with_pop, path = "C:/Users/lahebert.REDMOND/Documents/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/stops_summary_by_year_with_pop.csv")

```


Create some data sets: one with stop percentage by ethnicity for each division for all years. The second will have the same data broken out by year.  For each data set, add data with the ethnicity data by division populations for the neighborhood.  

```{r}
#Get the percentage of each stop type broken down by ethnicity
stop_type_by_ethnicity_citywide_all_years <- stops_data %>% 
  filter(!is.na(stop_type)) %>%
  group_by(stop_type, ethnicity) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(division = "CITY TOTAL")

#Add city-wide total 
stop_type_by_ethnicity_citywide_all_years <- 
  division_pop_gathered %>%
  filter(division == "CITY TOTAL", !is.na(ethnicity_percent)) %>%
  rename(stop_type = location) %>%
  bind_rows(stop_type_by_ethnicity_citywide_all_years)

stop_type_by_ethnicity_citywide_by_year <- stops_data %>% 
  filter(!is.na(stop_type)) %>%
  group_by(year, stop_type, ethnicity) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(division = "CITY TOTAL")


#Get the percentage of each ethnicity broken down by stop type
ethnicity_by_stop_type_citywide_all_years <- stops_data %>% 
  filter(!is.na(stop_type)) %>%
  group_by(ethnicity, stop_type) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(division = "CITY TOTAL")

write_csv(stop_type_by_ethnicity_citywide_all_years, path = "C:/Users/lahebert.REDMOND/Documents/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/stop_type_by_ethnicity_citywide_all_years.csv")

```