---
title: "Capstone Graphs"
author: "Lael Hebert"
date: "April 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
Load packages and data

```{r }
library(ggplot2)
library(readr)
library(dplyr)
library(tidyr)
stops_summary_all_years_with_pop <- read_csv("~/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/stops_summary_all_years_with_pop.csv")
stops_summary_by_year_with_pop <- read_csv("~/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/stops_summary_by_year_with_pop.csv", 
    col_types = cols(ethnicity_count = col_double()))
stop_type_by_ethnicity_citywide_all_years <- read_csv("~/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/stop_type_by_ethnicity_citywide_all_years.csv")
ethnicity_by_stop_type_divisions_by_year <- read_csv("~/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/ethnicity_by_stop_type_divisions_by_year.csv")
ethnicity_by_stop_type_citywide_all_years <- read_csv("~/Learning/Springboard/Introduction to Data Science/Capstone/Data/Output/ethnicity_by_stop_type_citywide_all_years.csv")

```

Start by graphing ethnicity_percent by populations and stops city-wide 
```{r, echo=FALSE}
stops_summary_all_years_with_pop %>% 
  filter(division == "CITY TOTAL",  ethnicity != "total_pop", location == "population") %>%
    ggplot(aes(x=ethnicity, y = ethnicity_percent)) + 
    geom_bar(aes(fill = location), stat = 'identity', position = 'dodge') +
    ggtitle("Ethnicity distributions")

stops_summary_all_years_with_pop %>% 
  filter(division == "CITY TOTAL",  ethnicity != "total_pop") %>%
    ggplot(aes(x=ethnicity, y = ethnicity_percent)) + 
    geom_bar(aes(fill = location), stat = 'identity', position = 'dodge') +
    ggtitle("Ethnicity distributions")

```

Now let's look at ethnic breakdowns by
Let's look at vehicular versus pedestrian stops.
```{r, echo=FALSE}
ggplot(stop_type_by_ethnicity_citywide_all_years, aes(x= ethnicity, y = ethnicity_percent)) + 
  geom_bar(aes(fill=stop_type), stat= 'identity', position='dodge') +
  ggtitle("Ethnicity by stop type")
ggplot(ethnicity_by_stop_type_citywide_all_years, aes(x= ethnicity, y = ethnicity_percent)) + 
  geom_bar(aes(fill=stop_type), stat= 'identity', position='dodge') +
  ggtitle("Stop type by ethnicity")

```

It looks like the black population has a lot more traffic stops than one would expect. Let's graph the percentage of stops that are for the black population by year
```{r, echo=FALSE}

stops_summary_by_year_with_pop %>% 
  filter(division == "CITY TOTAL", ethnicity == "BLACK", location == "stops", year != 2019) %>%
  ggplot(aes(x = year, y = ethnicity_percent)) + geom_point() 

```

Holy cow! It looks like the percentage of blacks in traffic stops is increasing. Let's add a regression line to the graph and look at the regression parameters.
```{r, echo=FALSE}
stops_summary_by_year_with_pop %>% 
  filter(division == "CITY TOTAL", ethnicity == "BLACK", location == "stops", year != 2019) %>%
  ggplot(aes(x = year, y = ethnicity_percent)) + geom_point() + 
  geom_smooth(method = 'lm', formula = y~x) +
  ggtitle("Black Stops")

summary(lm(ethnicity_percent ~ year, filter(stops_summary_by_year_with_pop, division == "CITY TOTAL", ethnicity == "BLACK", location == "stops")))

```
The percentage of stops of black people is increasing by about 0.8 percentage points a year. Let's look at each of the divisions
```{r, echo=FALSE}

#graph it:
ggplot(stops_summary_all_years_with_pop, aes(x=ethnicity, y = ethnicity_percent)) + geom_bar(aes(fill = location), stat = 'identity', position = 'dodge') + facet_grid(division ~.)

graph_district_distributions <- function(x) {
  stops_summary_all_years_with_pop %>% 
    filter(division == unique(stops_summary_all_years_with_pop$division)[x],  ethnicity != "total_pop") %>%
    ggplot(aes(x=ethnicity, y = ethnicity_percent)) + 
    geom_bar(aes(fill = location), stat = 'identity', position = 'dodge') +
    ggtitle(unique(stops_summary_all_years_with_pop$division)[x])
}
 
# for (i in 1:length(unique(stops_summary_all_years$division))) {
#   graph_district_distributions(i)
# }

graph_district_distributions(1)
graph_district_distributions(2)
graph_district_distributions(3)
graph_district_distributions(4)
graph_district_distributions(5)
graph_district_distributions(6)
graph_district_distributions(7)
graph_district_distributions(8)
graph_district_distributions(9)
graph_district_distributions(10)
graph_district_distributions(11)
graph_district_distributions(12)
graph_district_distributions(13)
graph_district_distributions(14)
graph_district_distributions(15)
graph_district_distributions(16)
graph_district_distributions(17)
graph_district_distributions(18)
graph_district_distributions(19)
graph_district_distributions(20)
graph_district_distributions(21)
graph_district_distributions(22)
```

##

Try some more graphs. Plot the black population verses the number of black stops for each division. Then plot the population percentage verses stops percentage by division for blacks.

```{r, echo=FALSE}
stops_summary_all_years_with_pop %>% 
  select(1,2,3,5) %>%
  filter(ethnicity == "BLACK", division != "CITY TOTAL") %>%
  spread(key = location, value = ethnicity_count) %>%
  ggplot(aes(x=population, y=stops)) + 
  geom_point() +
  geom_smooth(method = 'lm', formula = y~x) + 
  ggtitle("Black counts by division ")


stops_summary_all_years_with_pop %>% 
  select(1,2,4,5) %>%
  filter(ethnicity == "BLACK", division != "CITY TOTAL") %>%
  spread(key = location, value = ethnicity_percent) %>%
  ggplot(aes(x=population, y=stops)) + 
    geom_point() +
    geom_smooth(method = 'lm', formula = y~x) + 
  ggtitle("Black percentages by division ")

stops_summary_all_years_with_pop %>% 
  select(1,2,4,5) %>%
  filter(ethnicity == "BLACK", division != "CITY TOTAL") %>%
  spread(key = location, value = ethnicity_percent) %>%
  ggplot(aes(x=population, y=stops)) + 
    geom_point() +
    ylim(0,0.7) + xlim(0,.7) +
  geom_abline(slope = 1, intercept = 0, color = 'red') +
  ggtitle("Black percentages by division")


# Line of fairness graph: 
stops_summary_all_years_with_pop %>% 
    ggplot() + ylim(0,0.7) + xlim(0,.7) +
    geom_abline(slope = 1, intercept = 0, color = 'red') +
    ggtitle("Black percentages by division") +
    xlab("population") + ylab("stops")
```
Let's look at the percentages graph for all ethnicities
```{r, echo=FALSE}
stops_summary_all_years_with_pop %>% 
  select(1,2,4,5) %>%
  filter(ethnicity != "total_pop", division != "CITY TOTAL") %>%
  spread(key = location, value = ethnicity_percent) %>% 
  ggplot(aes(x=population, y= stops, col = ethnicity)) + 
  geom_point() + 
  ggtitle("Division Ethnicities Stops by Population Percentage")
```

Let's create a data set with only data for the black population. This will help us identify the points above the regression line.
```{r, echo=FALSE}
black_div_sum <- filter(stops_summary_all_years_with_pop, ethnicity == "BLACK")
black_div_sum$ethnicity = NULL
black_div_stops <- filter(black_div_sum, location == "stops")
black_div_pop <- filter(black_div_sum, location == "population")
black_div_sum <- left_join(black_div_pop, black_div_stops, by = "division") %>%
  rename(pop_count = ethnicity_count.x, pop_percent = ethnicity_percent.x, stop_count = ethnicity_count.y, stop_percent = ethnicity_percent.y) %>%
  mutate(stop_to_pop_count = stop_count / pop_count) %>%
  mutate(stop_to_pop_percent = stop_percent / pop_percent)
black_div_sum$location.x = NULL
black_div_sum$location.y = NULL
black_div_sum[1:22,]
```
Divisions to focus on:
+ Southeast - it's above both the percent and count trend lines
+ Southwest - it's below the percent line but above the count line due to larger ratio of stops to population
+ Central - it's on the percent line but way above the count line due to very large ratio of stops to population
+ Newton - it's way above the percent trend line also avove the  count trend lines
+ Hollywood -it's above both the percent and count trend lines
+ Olympic - it's above the percent trend line and has a huge stop to population ratio
Also, possibly:
+ West Los Angeles - it's above the percent trend line and also has a large white population



```{r, echo=FALSE}

```
