---
title: "LAPD data exploration"
author: "Lael Hebert"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r}
library(dplyr)
library(readr)
library(lubridate)
library(stringr)
library(tidyr)
```

## R Load data

Start by loading all the data into R with good column types. Fix column names.

```{r}
census_data <- read_csv("Census_Data_by_Neighborhood_Council.csv", 
    col_types = cols(Hispanic_pop = col_double(), 
        Owner_occ = col_skip(), Renter_occ = col_skip()))
census_data <- rename(census_data, neighborhood = NC_Name, Total_pop = "Total Population")

reporting_districts <- read_csv("LAPD_Reporting_Districts.csv", 
    col_types = cols(AREA = col_skip(), OBJECTID = col_skip(), 
        PERIMETER = col_skip(), PREC = col_character(), 
        REPDIST = col_character(), Shape__Area = col_skip(), 
        Shape__Length = col_skip()))
reporting_districts <- rename(reporting_districts, reporting_district = REPDIST, division_number = PREC, division = APREC, bureau = BUREAU, basic_car = BASICCAR)

stops_data <- read_csv("Vehicle_and_Pedestrian_Stop_Data_2010_to_Present.csv", 
    col_types = cols(`Form Reference Number` = col_character(), 
        `Officer 1 Serial Number` = col_character(), 
        `Reporting District` = col_character(), 
        `Stop Date` = col_date(format = "%m/%d/%Y"), 
        `Stop Number` = col_character(), 
        `Stop Time` = col_time(format = "%H:%M")))
names(stops_data) = c("stop_number", "form_reference_number", "gender_code", "ethnicity_code", "ethnicity", "date", "time", "officer_1_serial", "officer_1_division_num", "division_1", "officer_2_serial", "Officer_2_division_num", "division_2", "reporting_district", "stop_type", "post_stop_activity")

division_populations_all <- read_csv("Division_populations.csv")
division_pop <- Division_populations <- read_csv("Division_populations.csv", 
    col_types = cols(Asian_pop = col_skip(), 
        Black_pop = col_skip(), Hispanic_pop = col_skip(), 
        Other_pop = col_skip(), White_pop = col_skip(), 
        total_pop = col_skip()))
division_pop <- rename(division_pop, WHITE = white_percent, BLACK = black_percent, ASIAN = asian_percent, HISPANIC = hispanic_percent, OTHER = other_percent)


```
Clean up the stops data: 
Add a year column. 
If the reporting district starts with 0, remove the 0.
Join the stops data with the reporting districts data to get the division and bureau where the stop took place. 
Rename the new columns in the stops data.

```{r}
stops_data <- stops_data %>%
  mutate(year = year(date)) %>%
  mutate(reporting_district = ifelse(grepl(pattern = "^0", reporting_district), 
                                 str_sub(reporting_district, start = 2), reporting_district)) %>%
  left_join(reporting_districts, "reporting_district") %>%
  rename(stop_division_number = division_number, stop_division = division, stop_bureau = bureau, stop_basic_car = basic_car)


```

The population data is missing from the hispanic column in the census data. Fill it in by subtracting the popultions of the other ethnicities from the total population. Calculate the total population of the city. Then created a vector of the percentages of each ethnicity city-wide. 


```{r}
census_data$Hispanic_pop <- census_data$Total_pop - 
  census_data$White_pop -
  census_data$Black_pop -
  census_data$Ameri_es_pop -
  census_data$Asian_pop -
  census_data$Hawn_pi_pop -
  census_data$Other_pop -
  census_data$Multi_pop

city_pop_total <- sum(census_data$Total_pop)
city_wide_percentages <- census_data[,3:10] %>%
  summarise_each(funs(sum))/city_pop_total

city_pop_total
city_wide_percentages



```

According to the [Census.gov website](https://www.census.gov/quickfacts/fact/table/losangelescitycalifornia,losangelescountycalifornia/POP010210), Los Angelos had a population of 3.79 million people in 2010. This is very close to the 3.73 million reported in our census data set from the neighborhood council. The percentages by each ethnicity also match well. It looks like our calculation for the hispanic population will work.

Now, let's combine all ethnicities that represent less than 2% of the population as "other." We'll change the ethnicity column in stops_data but leave the ethnicity_code column alone in case we need the original values later. 
```{r}
unique(stops_data$ethnicity)
stops_data <- stops_data %>% 
  mutate(ethnicity = ifelse(ethnicity %in% c("OTHER", "AMERICAN INDIAN", "MULTI-DESCENTS"), "OTHER", ethnicity))
unique(stops_data$ethnicity)


```
Add a column to the census_data called All_other_pop with the sum of the populations of ethnic groups with less than 2% of the population.  
```{r}
census_data <- census_data %>% mutate(
  All_other_pop = 
    Ameri_es_pop +
    Hawn_pi_pop +
    Other_pop + 
    Multi_pop
)
```
Subset to just 2010 and find percentages of each ethnicity by stop division
```{r}


stops_summary_2010 <- stops_data %>% 
  filter(year == 2010, !is.na(stop_division)) %>%
  group_by(stop_division, ethnicity) %>%
  summarise(ethnicity_count = n()) %>%
  mutate(ethnicity_percent = ethnicity_count/sum(ethnicity_count)) %>%
  mutate(location = "stops") %>%
  rename(division = stop_division)

stops_summary_2010$ethnicity_count = NULL

#reshape division_pop so that is in the same form as stops_summary
division_pop <- division_pop %>%
  gather("ethnicity", "ethnicity_percent", 2:6) %>%
  mutate(location = "population")

stops_summary_2010_with_pop <- bind_rows(stops_summary_2010, division_pop)

#graph it:
ggplot(stops_summary_2010_with_pop, aes(x=ethnicity, y = ethnicity_percent)) + geom_bar(aes(fill = location), stat = 'identity', position = 'dodge') + facet_grid(division ~.)


#join stops_summary_2010 to division population data
#reshape stops_summary_2010 with gather() so that it has these columns:
  #stop_division, ethnicity, location, ethnicity percentage
  #location column will contain "stops" or "neighborhood" 


#scratchpad stuff I might use  later
#stops_data %>% group_by(gender_code) %>% summarise(totals = n()/nrow(stops_data))
#stops_data %>% group_by(ethnicity_code) %>% summarise(totals = n()/nrow(stops_data))
#gather(stops_hollywood_2010, "location", "ethnicty_percentage", 4:5)
#ggplot(hollywood_2010_summary, aes(x=ethnicity, y = ethnicty_percentage)) + geom_bar(aes(fill = location), stat = 'identity', position = 'dodge')
```

